---
title: "TP 1 - Introduction à `R markdown`"
author: "Paul Bastide, Chloé Serre-Combe, Alexandre Capel"
date: "2024-2025"
output: html_document
---

## `R Markdown`

`R Markdown` allows you to create dynamic documents that include formatted text, equations, and `R` code.

This tool is very useful for writing technical reports. It enables you to present the context of the problem, the method of resolution, and the results of the analysis all in a single document.

Many resources are available online.
See for example  
[the official introduction](https://rmarkdown.rstudio.com/lesson-1.html)
to `R markdown`, and the [sheets](https://raw.githubusercontent.com/rstudio/cheatsheets/master/rmarkdown-2.0.pdf).

## Installation

To use `R markdown`, you just have to install the library by writing the following command in the console:

```{r install-rmd, eval = FALSE}
install.packages("rmarkdown")
```


## Document creation

Once the installation done, the easier way is to create a new document with the `RStudio` interface .

Select in the menu :
```
File > New File > R markdown...
```

In the tab `Document` (default), you can write the title and the author's name of the document;

There arer three formats:

* `HTML`: allows for dynamic formatting, useful for websites.

* `PDF`:  allows for fixed formatting, suitable for an "official" or printed output.

* `Word`:allows for the generation of editable documents that can be shared with collaborators.

For this session, we choose the `HTML` format, as it is the simplest for layout purposes.

## First document
Once the document is created, `RStudio` provides default "educational" content.

You can compile this document by clicking on the `Knit` command at the top left, next to a yarn ball.

If everything works correctly, `RStudio` will generate an `HTML` document and open it in a new window.

You can study this initial document and see how the source (the `.Rmd`) influences the output (the `.html`)


## R Code and Graphs

Let’s suppose we conduct the following experiment.  
A urn contains 300 balls: 100 red, 100 blue, and 100 green.  
We draw a ball at random, note its color, and set it aside.  
We repeat this experiment 60 times.  

With `R`, we can simulate this random experiment as follows.

```{r balls-experiment}
set.seed(1289)                         ## Reproducibility

urn <- c(rep("red", 100),              ## 100 red balls
          rep("blue", 100),             ## 100 blue balls
          rep("green", 100))            ## 100 green balls 

n_exp <- 60                            ## Number of times the experiment is conducted

sampled_balls <- sample(urn,          ## sampling the balls
                        n_exp,         ## n_exp times
                        replace = FALSE) ## without replacement
```

We can then display a summary of the experiment (the number of times each ball was drawn).

```{r balls-results}
res <- table(sampled_balls)            ## Summary of the sample
res
```

And plot this result.

```{r balls-results-plot}
barplot(res)                           ## Bar plot
```

It is also possible to use the results in the text.

For example, from the `r n_exp` experiments, we drew the blue ball `r res[1]` times.

## Exercise

> **To do**: A *compiled* report in `html` or `pdf` format.

Write an `R Markdown` document that addresses the problem outlined below.  
The document should be self-sufficient and clearly present the problem, the analyses, and the conclusions.

It should contain sections, words highlighted in **bold** and *italic*, and `code` in the text.

It will also include `R` code blocks that will be used to address the problem.

It may contain $\LaTeX$ equations, and numbered or unnumbered lists.

---

> We are fishing for fish in the Lez.
> 
> We assume there are a total of $10,000$ fish in the Lez, of which $2,000$ are red, $3,000$ are green, and $5,000$ are blue
> (the Lez is very polluted).
> However, we assume that we do not have access to this information
> (we do not know how many fish there are in total, nor how many of each color).
>
> We ask the question: "What is the proportion of red, green, and blue fish in the Lez?"
> 
> We catch $100$ fish, which we keep to display in an aquarium, and we note their color.
>
> Describe the statistical experiment: the question posed, individuals, population, sample, size, and type of variable measured.
> Does the variable have incompatible, exhaustive, and unambiguous categories? 
>
> Simulate the data corresponding to this experiment using the `sample` function.
>
> Calculate the empirical frequency of each color of fish using the `table` function.
> What is the estimated proportion of red, green, and blue fish?
> Does this result surprise you?
>
> Plot these empirical frequencies using the `barplot` function.
>
> What happens when you compile your document multiple times in a row?
> Do the numerical results change? 
> Is this normal?
> Use the `set.seed` function to produce a reproducible document.
>
> We assume that, due to budget constraints, we can only catch $10$ fish.
> What results do you obtain? Is the estimation of proportions accurate in this case?
>
> Now assume that, thanks to an online fundraising campaign, we can catch $1,000$ fish. Does this improve the estimation?

--- 
