---
title: "IG4 - Visualisation"
subtitle: "Introduction"
author: "Paul Bastide, Chloé Serre-Combe, Alexandre Capel"
output:
  ioslides_presentation:
    fig_width: 7
    fig_height: 4
  self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- ************************************************************************ -->
# Some informations
<!-- ************************************************************************ -->

## Teachers

- Paul Bastide
    - paul.bastide@umontpellier.fr
    - Applied statistics, evolution

- Chloé Serre-Combe
    - chloe.serre-combe@umontpellier.fr
    - Extreme value theory, spatial statistics

- Alexandre Capel
    - alexandre.capel@umontpellier.fr
    - Extreme value theory, graphical model

## Module

* Aim 
  * how to summarize a data frame ?
  * especially : how to make the reading of the data easier ?

* Plan 
  * `Rmarkdown` + `ggplot2` (course 1)  
  * Visualisation quali/quanti (course 2 + 3)
  * `shiny` (course 4)

## Exams

I don't know yet, maybe a small practical work or MCQ...

## Data

* [Open Data Montpellier Méditerranée Métropole](https://data.montpellier3m.fr/datasets)

* [data.gouv.fr](https://www.data.gouv.fr/fr/),
with its [Twitter account](https://twitter.com/datagouvfr)

* [data.europa.eu](https://data.europa.eu/euodp/en/home) and the [European Data Portal](https://www.europeandataportal.eu/en)

* [data.sncf.com](https://data.sncf.com)

* [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday#datasets), on Twitter [#tidytuesday](https://twitter.com/search?q=%23tidytuesday)

* [Five Thirty Eight](https://data.fivethirtyeight.com/)
and its associated [package](https://CRAN.R-project.org/package=fivethirtyeight) 

* [KD Nuggets](https://www.kdnuggets.com/datasets/index.html)

* ...

## Ressources

* *Fundamentals of Data Visualization*, Claus O. Wilke:
  * [Book](https://clauswilke.com/dataviz/)
  * [Course](https://wilkelab.org/SDS375/)

* *Data Visualization, A practical introduction*, Kieran Healy:
  * [Book](https://socviz.co/)

* Page [GitHub](https://github.com/alxkpl/ig4-dataviz) for the course

## GitHub `ig4-dataviz`

* Based on [repo GitHub of Paul Bastide](https://github.com/pbastide/do3-dataviz) and [de Clhoé Serre-Combe](https://github.com/chloesrcb/ig4-dataviz)

* [Mine](https://github.com/alxkpl/ig4-dataviz)

* The slides are in the folder `slide`

* The practical works are in the folder `TP`

<!-- ************************************************************************ -->
# R Markdown
<!-- ************************************************************************ -->

## `Markdown` :  lightweight markup language 

**Input** :

```
I can write in *italic*, in **bold**, and even `code`.

Thanks to $\LaTeX$, I kwnow the Euler's identity : $e^{i\pi} + 1 = 0$.
```

**Output** :

I can write in *italic*, in **bold**, and even  `code`.

Thanks to  $\LaTeX$,  I kwnow the Euler's identity : $e^{i\pi} + 1 = 0$.


$~$
**Format** :
PDF, HTML, documents, slides, ...

## `R Markdown` : Reproducible code

**Input** :

````
The function `rnorm()` simulate gaussian variables.
`r ''````{r}
set.seed(12890926) # seed (Quia Sapientia)
rnorm(5)           # 5 standard gaussian variables
`r ''````
````

**Output** :

The function `rnorm()` simulate gaussian variables.
```{r}
set.seed(12890926) # seed (Quia Sapientia)
rnorm(5)           # 5 standard gaussian variables
```


## `R Markdown` : Reproducible graphics 

The Maunga Whau Volcano (Mt Eden), Auckland, Nouvelle-Zélande

```{r demo}
filled.contour(volcano, color.palette = terrain.colors, asp = 1)
```

## `R Markdown`

* [Full introduction](https://rmarkdown.rstudio.com/lesson-1.html)

* [Sheets](https://raw.githubusercontent.com/rstudio/cheatsheets/master/rmarkdown-2.0.pdf)

* [RStudio](https://rstudio.com/)

* Theses slides on [GitHub](https://github.com/alxkpl/ig4-dataviz)
