---
title: "TP 1.2 - Introduction to `R Markdown` and `ggplot2`"
author: "Paul Bastide, Chloé Serre-Combe, Alexandre Capel"
date: "2024-2025"
output: html_document
---

## Setup

Clone or download the course repository from `GitHub`.

This folder contains a `.Rproj` file, which defines a "project" in RStudio.  
The project allows you to set a number of default behaviors and a folder structure.

Open this project in `RStudio`.

Create a new `.Rmd` document with a `pdf` output format.

To compile `pdf` documents, you will need LaTeX.  
If you don't have it installed on your machine yet,  
you can download the full (and large) LaTeX distribution suitable for your machine using `apt-get` or another package manager.  
Alternatively, you can install a minimal version, which is sufficient for compiling documents in `pdf` format, by installing the [`tinytex`](https://yihui.org/tinytex/) package:

```{r install-rmd, eval = FALSE}
install.packages('tinytex')
tinytex::install_tinytex()
```

You will complete this document progressively by following the text’s instructions.  
The points to write are indicated by the tag:

> **To do**: Create a new `.Rmd` document to compile a `pdf`.

---

## `ggplot2`

`ggplot2` is an implementation of the  
["grammar of graphics"](https://en.wikipedia.org/wiki/Ggplot2),  
allowing you to create precise and informative scientific figures easily in `R`.

Many resources are available online.  
See for example the [official documentation](https://ggplot2.tidyverse.org/index.html)  
as well as the [cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf).

## Installation

To use `ggplot2`, simply install the associated package by typing the following command in the console:

```{r install-ggplot2, eval = FALSE}
install.packages("ggplot2")
```

Then load the package with the command:

```{r load-ggplot2}
library(ggplot2)
```

## Data: Average Temperatures in Montpellier

We will use the same temperature data as in the course.

Thanks to the project structure, we can easily reference it using the `here` package.

```{r install-here, eval = FALSE}
install.packages("here")
```

```{r load-here}
library(here)
```

We load the data using the `read.csv` function.

```{r temperatures}
temps_average <- read.csv(file = here("data", "temp_average.csv"))
```

We will keep only the temperatures measured in Montpellier.  
To do this, we use the `subset` function.

```{r temperatures-montpellier}
temps_montpellier <- subset(temps_average, city == "Montpellier")
```

We can display the first rows of the data table with the `head` function.

```{r temp-head}
head(temps_montpellier)
```

> **To do**: Load and present the data in your document.  
Briefly describe the structure of the data table (dimensions, meaning of variables, possible values, units, ...).  
Mention the source and the time period considered (see slides).

## Data: Basic Statistics

### `which.min`

The `which.min` function allows you to find the minimum of a vector:

```{r which-min}
x <- c(2, 3, 5, 1, 8)
min(x)
which.min(x)
x[which.min(x)]
```

### `as.Date`

The `as.Date` function converts a day into a date:

```{r as-Date}
as.Date(1, origin = "2019-12-31")           ## January 1, 2020
as.Date(31, origin = "2019-12-31")          ## January 31, 2020
as.Date(31 + 29, origin = "2019-12-31")     ## February 29 (2020 is a leap year)
as.Date(366, origin = "2019-12-31")         ## December 31 (2020 is a leap year)
```

> **To do**: What is the average temperature over the year in Montpellier?  
On which day is it, on average, the hottest in Montpellier?  
The coldest?  

You can use the `mean`, `which.min`, `which.max`, and `as.Date` functions.  

**Note**: The "average" year has 366 days to account for leap years.

## Temperatures Represented on `y`

We want to graphically represent the average temperatures for each day of the year.

First, we want to plot:

* the `temperature` column on `y` (ordinates)

* the `day_of_year` column on `x` (abscissa).

We use the `ggplot` function following this skeleton:

```{r temperatures-plot-y, eval = FALSE}
ggplot(---,
       aes(x = ---,              
           y = ---)) +              
  geom_---()                     
```

The `ggplot` function takes the data as input and then the "aesthetics" given by `aes`.

The functions `geom_---` indicate what to plot.  
You can try using `geom_line` or `geom_point` here.

> **To do**:  
Plot the average temperature as a function of the day of the year in Montpellier.  
Vary the type of graph and their size (lines or points, larger or smaller), as well as their color.

## Multiple Cities

Now we return to the complete dataset, which includes average temperatures for three cities: Embrun, Montpellier, and Lille.

```{r temperatures-head}
head(temps_average)
```

We want to represent the three curves on the same graph.

To do this, we need to distinguish the cities using a distinct graphic symbol.

For example, we can differentiate the cities by their color.  
To do this, simply add the `color` parameter in the `aes`.

```{r temperatures-plot-y-color, eval = FALSE}
ggplot(temps_average,
       aes(x = ---,              
           y = ---,
           color = ---)) +              
  geom_---()                     
```

It is also possible to distinguish the cities by the type of line drawn.  
To do this, simply add the `linetype` parameter in the `aes`.

```{r temperatures-plot-y-line, eval = FALSE}
ggplot(temps_average,
       aes(x = ---,              
           y = ---,
           linetype = ---)) +              
  geom_line()                     
```

> **To do**:  
Plot the average temperature as a function of the day of the year in the three cities, using either color or line type to distinguish them.

## Temperatures Represented in `color`

You can change the assignments in the `aes` as desired, depending on the desired effect.  
For example, you might want to represent temperatures with *different colors*, instead of representing them on the y-axis.

To do this, simply swap `y` and `color` in the previous graph.

```{r temperatures-plot-color, eval = FALSE}
ggplot(temps_average,
       aes(x = ---,              
           y = ---,
           color = temperature)) +              
  geom_point(size = 5)                     
```

> **To do**:  
Plot the average temperature as a function of the day of the year in the three cities, using color to represent the temperature.  
Vary the assignment of different data columns to the different `aes` and observe the result.

## Mixes

You can combine the different components already presented to obtain new graphs.

> **To do**:  
For example, try to achieve the following representations:
> 
> * Temperatures on `y`, with a **different** color and line type for each city.
> 
> * Temperatures on `y`, a different line type for each city, and a color for each month.
> 
> * Temperatures in `color`, represented vertically.
> 
> * Graph showing the month as a function of the day of the year.

## Formatting

By default, the background is gray.  
You can change the appearance by using a different "theme."  
See for example `theme_light` or `theme_dark`.

You can also change the axis names using the `xlab` and `ylab` functions.

The `ggtitle` function allows you to add a title to the graph.

You simply add these options using the `+` operator:

```{r temperatures-plot-y-theme, eval = FALSE}
ggplot(---, aes(---)) +              
  geom_---() +
  xlab(---) + ylab(---) + ggtitle(---) + 
  theme_---()
```

> **To do**:  
Take your favorite graph and modify it to make it "publishable":  
add a theme of your choice, name the axes informatively (with units), and add a title.  
Provide the statistical formulation associated with the chosen graph (question, individual, population, sample, measured variable).
